---
title: "EDS 211 parallel test"
output: html_document
date: "2023-02-06"
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse) ### includes purrr
library(parallel)
library(furrr)
library(raster)
library(tictoc)
```

### do a slow thing?

```{r}
rast_fs <- list.files('data', full.names = TRUE)

r <- raster::raster(rast_fs[1])

plot(r)

df <- as.data.frame(r) %>%
  drop_na() %>%
  mutate(type = case_when(x < 0.1 ~ 'low',
                          x < 0.25 ~ 'medium',
                          x < 0.5 ~ 'high',
                          TRUE ~ 'wow'))
summary_df <- df %>%
  group_by(type) %>%
  summarize(n_cells = n())
```

### in a loop

```{r}
system.time({
out_list_loop <- vector('list', length = length(rast_fs)) %>%
  setNames(rast_fs)

for(f in rast_fs) {
  r <- raster::raster(f)
  
  df <- as.data.frame(r) %>%
    setNames('x') %>%
    drop_na() %>%
    mutate(type = case_when(x < 0.1 ~ 'low',
                            x < 0.25 ~ 'medium',
                            x < 0.5 ~ 'high',
                            TRUE ~ 'wow'))
  summary_df <- df %>%
    group_by(type) %>%
    summarize(n_cells = n(),
              file = basename(f))
  
  out_list_loop[[f]] <- summary_df
}
}) ### 20-30 seconds

out_df_loop <- out_list_loop %>%
  bind_rows()

```

### Create a function

```{r}
process_rast_file <- function(f) {
  r <- raster::raster(f)
  
  df <- as.data.frame(r) %>%
    setNames('x') %>%
    drop_na() %>%
    mutate(type = case_when(x < 0.1 ~ 'low',
                            x < 0.25 ~ 'medium',
                            x < 0.5 ~ 'high',
                            TRUE ~ 'wow'))
  summary_df <- df %>%
    group_by(type) %>%
    summarize(n_cells = n(),
              file = basename(f))
  
  return(summary_df)
}

boring <- function(t = 3) {
  Sys.sleep(t)
  return(t)
}

```

### lapply

Iterates over a sequence (vector, list, dataframe columns, etc) and applies some function.  Returns results in a list (very flexible)

```{r}
ptm <- proc.time()
out_list_lapply <- lapply(X = rast_fs, FUN = process_rast_file)
proc.time() - ptm ### elapsed ~ 21-23 sec

out_df_lapply <- bind_rows(out_list_lapply)

t_vec <- 1:10
ptm <- proc.time()
t_list <- lapply(t_vec, boring)
proc.time() - ptm ### elapsed 55 sec

```

### parallel::mclapply

Iterates over a sequence (vector, list, dataframe columns, etc) and applies some function.  Returns results in a list (very flexible).

NOTE: have ONE person in each group run the code, then the next person!

```{r}
parallel::detectCores() ### 64!

ptm <- proc.time()
out_list_mclapply <- parallel::mclapply(X = rast_fs, FUN = process_rast_file, mc.cores = 4)
proc.time() - ptm ### elapsed ~ 10 sec with 4 cores, 3-5 sec with 15 cores

out_df_mclapply <- bind_rows(out_list_mclapply)
```

``` {r}
tic()
t_list <- mclapply(t_vec, FUN = boring, mc.cores = 5)
toc() ### elapsed ~ 15 sec

tvec_flip <- c(1:5, 10:6)
tic()
t_list <- mclapply(tvec_flip, boring, mc.cores = 5)
toc() ### elapsed ~ 11 sec

```

### purrr::map() version 1

```{r}
tic()
out_list_purrr <- purrr::map(rast_fs, process_rast_file)
toc() ### elapsed ~ 21-23 sec

out_df_purrr <- bind_rows(out_list_purrr)

```

```{r}
tic()
out_df_purrr2 <- purrr::map_dfr(rast_fs, process_rast_file)
toc() ### elapsed ~ 21-23 sec
```


### furrr::future_map() - note version 0.3.1, still well in development

```{r}

plan(multisession, workers = 4)
tic()
out_list_furrr <- furrr::future_map(rast_fs, process_rast_file)
toc() ### elapsed ~ 28 sec

out_df_purrr <- bind_rows(out_list_purrr)

```


